---
title: "assignment-4"
author: "Gage Clawson"
date: "11/13/2018"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(vcdExtra)
library(car)
library(effsize)
library(ggsignif)
library(ggpubr)
library(knitr)
library(FSA)
library(kableExtra)
library(xtable)
library(kableExtra)
```

```{r}
# read in the lobster traps dataset
lob_traps <- as.data.frame(read_csv("lobster_traps.csv"))


```

```{r}
# Read in the lobster size data set and make it tidy

lob_size <- read_csv("lobster_size_abundance.csv")
lob_size1 <- as.data.frame(lob_size)

lob_size2 <- expand.dft(lob_size1, freq = "COUNT") ##expand.dft filters out all count = 0 rows and makes data tidy (each observation has it's own row)
```
Yes boss

```{r}
lob_size3 <- mutate(lob_size2, DATE = as.Date(DATE, format = '%d-%b-%y')) # make date formats same for both datasets yyyy-mm-dd

three_sites <- c("AQUE", "CARP", "MOHK")

lob_traps1 <- mutate(lob_traps, DATE = as.Date(DATE, format = '%m/%d/%y')) %>%
  filter(SITE %in% three_sites, TRAPS != 0) %>%
    expand.dft(freq = "TRAPS") 
  


# make date formats the same and filter for only the 5 sites 

```




4. Proportions of “legal” lobsters at the 5 sites in 2017
The legal minimum carapace size for lobster is 82.6 mm. What proportion of observed lobsters at each site are above the legal minimum? Does that proportion differ significantly across the 5 sites? Note: We’ll be doing chi-square in labs next week, or go ahead with maximum resourcefulness and check out the chisq.test() function on your own!

```{r}
# filter for 2017
# filter for above 82.6 mm

legal_min1 <- lob_size3 %>%
  filter(YEAR == 2017) %>%
  mutate(above_legal = case_when(
    SIZE >= 82.6 ~ 'yes',
    SIZE < 82.6 ~ 'no'
  )) %>%
  count(SITE, above_legal) 

total <- c(sum(legal_min1$n[legal_min1$SITE == "AQUE"]), sum(legal_min1$n[legal_min1$SITE == "AQUE"]), sum(legal_min1$n[legal_min1$SITE == "CARP"]),sum(legal_min1$n[legal_min1$SITE == "CARP"]), sum(legal_min1$n[legal_min1$SITE == "IVEE"]),sum(legal_min1$n[legal_min1$SITE == "IVEE"]), sum(legal_min1$n[legal_min1$SITE == "MOHK"]),sum(legal_min1$n[legal_min1$SITE == "MOHK"]), sum(legal_min1$n[legal_min1$SITE == "NAPL"]),sum(legal_min1$n[legal_min1$SITE == "NAPL"]))

totaln <- c(51, 16, 526, 179, 476, 130,154, 24, 75, 37)


proportions <- totaln/total

legal_min1["proportions"] <- proportions

legal_min2 <- legal_min1 %>%
  filter(above_legal == "yes")

total_only_5_columns <- c(sum(legal_min1$n[legal_min1$SITE == "AQUE"]), sum(legal_min1$n[legal_min1$SITE == "CARP"]), sum(legal_min1$n[legal_min1$SITE == "IVEE"]), sum(legal_min1$n[legal_min1$SITE == "MOHK"]), sum(legal_min1$n[legal_min1$SITE == "NAPL"]))

legal_min2["Total_Lobsters"] <- total_only_5_columns


legal_min_final_table <- legal_min2 %>% mutate(props = round(proportions,2)) %>%
  select(SITE, n, Total_Lobsters, props)


#legal_min_final_table$SITE[legal_min_final_table$SITE=="AQUE"] <- "Arroyo Quemado"

kable(legal_min_final_table, "latex", col.names = c(" ", ' "Legal" Lobsters', "Total Lobsters", 'Proportion "Legal" Lobsters'), booktabs = T, align = "c") %>%
  kable_styling(latex_options = c( "striped"), full_width = FALSE) %>%
  row_spec(0, bold = T) %>%
  column_spec(1, bold = T) %>%
  add_header_above(c(" ", '"Legal" California Spiny Lobsters' = 3), bold = T)





```


Based on the obervations from five Long-Term Ecological Research (LTER) Sites in the Santa Barbara Channel:Arroyo Quemado (n= 67) , Naples Reef (n= 112 ), Mohawk Reef (n= 178), Isla Vista (n= 606), Carpinteria(n= 705 ),the proportion of observed lobsters that are above the legal minimum carapace size does not differ significantly by site.( *$\chi$$^2$*(4)= 0.11095, *p*= 0.9985)








